---
title: "fifa_wages"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r libraries}
library(ggplot2)
```

```{r import_data}
player_data <- read.csv("../fifa2019wages/FifaTrainNew.csv", stringsAsFactors = FALSE)
head(player_data)


for(i in 1:length(player_data$Club)) {
    if(is.na(player_data$Club[i]) | player_data$Club[i] == "NA"){
        player_data$Club[i] <- "MISSING"
    }
}

for(i in 1:length(player_data$International.Reputation)) {
    if(is.na(player_data$International.Reputation[i])){
        player_data$International.Reputation[i] <- 1
    }
}
#player_data$Club

(sum(is.na(player_data$Overall)))
sum(is.na(player_data$International.Reputation))
sum(is.na(player_data$Age))
sum(is.na(player_data$Club))

#ggplot(data = player_data, aes(x = Overall, y = WageNew)) + geom_point()

#ggplot(data = player_data, aes(x = Age, y = WageNew)) + geom_point()
#ggplot(data = player_data, aes(x = Finishing, y = WageNew)) + geom_point()

#player_data <- na.omit(player_data)
#(sum(is.na(player_data)))

#ggplot(data = player_data, aes(x = Potential*Age, y = WageNew)) + geom_point()
#ggplot(data = player_data, aes(x = Weak.Foot, y = WageNew)) + geom_point()
#ggplot(data = player_data, aes(x = International.Reputation, y = WageNew)) + geom_point()
#ggplot(data = player_data, aes(x = Real.Face, y = WageNew)) + geom_point()

#player_data <- na.omit(player_data)


#ggplot(data = player_data, aes(x = seq_len(nrow(player_data)), y = WageNew)) + geom_point()


#ggplot(data = player_data[player_data$Position == "GK",], aes(x = GKDiving, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "GK",], aes(x = GKPositioning, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "GK",], aes(x = GKKicking, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "GK",], aes(x = GKHandling, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "GK",], aes(x = GKReflexes, y = WageNew)) + geom_point()

#ggplot(data = player_data[player_data$Position == "ST",], aes(x = Finishing, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "CF",], aes(x = Finishing, y = WageNew)) + geom_point()
#ggplot(data = player_data[player_data$Position == "CAM",], aes(x = Finishing, y = WageNew)) + geom_point()

#levels(player_data$Position )

#May need to use diff stats for  GK?
#pos_vec <- c()
#for(elem in player_data$Position) {
#    if(elem == "CB" | elem == "LB" | elem == "LCB" | elem == "RB" | elem == "RCB" | elem == "LWB" | elem == "RWB"){
#        pos_vec <- c(pos_vec, "B")
#    } else if (elem == "CDM" | elem == "CM" | elem == "LCM" | elem == "LDM" | elem == "LM" |  elem == "LW" | elem == #"RCM" | elem == "RDM" | elem == "RW" | elem == "RM") {
#        pos_vec <- c(pos_vec, "M")
#    } else if (elem == "CAM" | elem == "CF" | elem == "LAM" | elem == "LF" | elem == "LS" |  elem == "RAM" | elem == #"RF" | elem == "RS" | elem == "ST") {
#        pos_vec <- c(pos_vec, "F")
#    }
#}



```


```{r import_test_data}
library(assertthat)
test_data <- read.csv("../fifa2019wages/FifaNoY.csv")
player_data <- read.csv("../fifa2019wages/FifaTrainNew.csv", stringsAsFactors = FALSE)

levels(player_data$Position)
cor_data <- na.omit(player_data)
#head(test_data)
#ggplot(data = player_data, aes(x = seq_len(nrow(player_data)), y = WageNew)) + geom_point()
#player_data <- player_data[is.na(player_data)] <- "0"
#player_data[is.na(player_data)] <- 0
cor_goalie <- subset(player_data, Position == "GK")
vals <- cor(cor_data[sapply(cor_data, is.numeric)])
wage_corr <- vals[,'WageNew']
wage_corr[order(wage_corr)]

vals_g <- cor(cor_goalie[sapply(cor_goalie, is.numeric)])
wage_corr_g <- vals_g[,'WageNew']
wage_corr_g[order(wage_corr_g)]


```

```{r sample_sol}

sol1 <- read.csv("SampleSol.csv")
sol2 <- read.csv("SampleSol2.csv")
FifaNoY <- read.csv("FifaNoy.csv")
head(FifaNoY)
head(sol1)
head(sol2)

over_rating_pred2<- lm(player_data$Wage ~ player_data$Overall)

over_rating_pred1 <- lm(player_data$Wage ~ player_data$Overall + player_data$International.Reputation + player_data$Potential + player_data$Reactions + player_data$Composure)
summary(over_rating_pred1)
anova(over_rating_pred1)

player_data$WageNew[order(player_data$WageNew)]

#player_data$HL <- ifelse(player_data$Overall > 85, "H", "L")

```


```{r init_pred}


over_rating_pred <- lm(player_data$Wage ~ player_data$Overall + player_data$International.Reputation + player_data$HL + player_data$Potential)


anova(over_rating_pred)

over_rating_pred1 <- lm(player_data$Wage ~ 0 + player_data$Overall + player_data$International.Reputation + player_data$Potential + player_data$Reactions + player_data$Composure)


anova(over_rating_pred1)
outliers <- player_data$WageNew[abs(rstandard(over_rating_pred1))>=2]


wages <- player_data$WageNew[order(player_data$WageNew)]


player_data_outliers <- player_data[order(player_data$WageNew)[1:9500], ]

#player_data_outliers

over_rating_pred2 <- lm(player_data_outliers$Wage ~ 0 + player_data_outliers$Overall + player_data_outliers$International.Reputation + player_data_outliers$Potential + player_data_outliers$Reactions + player_data_outliers$Composure)
 
anova(over_rating_pred2)
#over_rating_pred <- lm(player_data$Wage ~ 0 + player_data$Overall + player_data$International.Reputation + player_data$Potential + player_data$Reactions + player_data$Composure + player_data$Club)


#inverseResponsePlot(over_rating_pred)
over_rating_pred <- lm((player_data$Wage)^0.8 ~ 0 + player_data$Overall + player_data$International.Reputation + player_data$HL + player_data$Potential + player_data$Composure + player_data$Club)




anova(over_rating_pred)
inverseResponsePlot(over_rating_pred)
over_int <- over_rating_pred$coefficients[1]
over_slope <- over_rating_pred$coefficients[2]
over_slope <- over_rating_pred$coefficients[2]


#plot(over_rating_pred)

pred_wages <- predict(over_rating_pred, test_data)


plot(pred_wages)
result_matrix <- matrix(test_data$Ob)
result_matrix <- cbind(result_matrix, pred_wages)
colnames(result_matrix) <- c("Ob", "WageNew")
write.csv(result_matrix, "Test Regression", row.names = FALSE)
nrow(result_matrix)
```


```{r clean}

#ifelse(player_data == "", NA, player_dat)
sum(player_data == "")

wages <- player_data$WageNew[order(player_data$WageNew)]

order(wages)

#(table(new_w <- cut(wages, 20)))
#plot(table(new_w), ylim = c(0, 20), xlim = c(0, 10000))

hist(wages, xlim = c(0, 12000), ylim = c(0, 4000))

plot(wages)

length(wages)
length(which(wages > 2000))
```

```{r test}

train_func <- function(train_ratio, FUN, y_power = 1, response, ...) {
    size <- length(response)
    size_vec <- seq_len(size)
    train_rows <- sample(1:size, size = train_ratio * size, replace = FALSE)
    test_rows <- size_vec[-train_rows]

    predictors <- list(...)   
    
    lin_reg <- FUN(response[train_rows] ~ predictors[[1]][train_rows] + predictors[[2]][train_rows] + predictors[[3]][train_rows])# + predictors[[4]][train_rows])

    
    (coef <- lin_reg$coefficients)

    (error <- sum((response[test_rows] - (coef[1] + coef[2] * predictors[[1]][test_rows] + coef[3] * predictors[[2]][test_rows] + coef[4] * predictors[[3]][test_rows]))^2))

        
}

#average_rsq <- 0

times <- 50
for(i in seq_len(times)) {
    average_rsq <- average_rsq + train_func(0.9, lm, y_power = 1, player_data$WageNew, player_data$Age, player_data$Overall, player_data$International.Reputation) / times
}
formatC(average_rsq, format = "e", digits = 5)
train_func(0.9, lm, y_power = 1, player_data$WageNew, player_data$Age, player_data$Overall, player_data$International.Reputation)


```

```{r test}
tr_func <- function(train_ratio, y_power = 1, response, ...) {
    avg <- 0
    for(i in 1:10) {
        size <- length(response)
        size_vec <- seq_len(size)
        train_rows <- sample(1:size, size = train_ratio * size, replace = FALSE)
        test_rows <- size_vec[-train_rows]
    
        predictors <- list(...)   
        
        lin_reg <- lm(log((response[train_rows])^y_power) ~ 0 + predictors[[1]][train_rows] + predictors[[2]][train_rows] + predictors[[5]][train_rows] + predictors[[6]][train_rows])
        
        pred_wages <- predict(lin_reg, player_data[test_rows,])
        
      
        SSR <- sum((pred_wages - player_data$WageNew[test_rows])^2)
        avg <- avg + SSR / 10

    }
    avg
}

inverseResponsePlot(lm(log(WageNew) ~ Overall + Potential + new_pos, data = player_data))

tr_func(0.8, 1, player_data$WageNew, player_data$Overall, player_data$Potential, player_data$HL, player_data$International.Reputation, player_data$new_pos, player_data$HL )



```


```{r pred}
library("car")
test_data <- read.csv("FifaNoY.csv", stringsAsFactors = FALSE)

for(i in 1:length(test_data$Club)) {
    if(is.na(test_data$Club[i]) | !is.element(test_data$Club[i], teams) ){
        test_data$Club[i] <- "MISSING"
    }
    if(is.na(test_data$Overall[i])){
        test_data$Overall[i] <- median(test_data$Overall[i])
    }
    if(is.na(test_data$Age[i])){
        test_data$Age[i] <- median(test_data$Age[i])
    }
    if(is.na(test_data$International.Reputation[i])){
        test_data$International.Reputation[i] <- 1
    }
}

predi <- lm(log(player_data$WageNew) ~ 0 + player_data$Overall + player_data$Age + player_data$Club)

#pos_vec <- c()
#for(elem in test_data$Position) {
#    if (is.na(elem)) {
#        pos_vec <- c(pos_vec, "0")
#    } else if(elem == "CB" | elem == "LB" | elem == "LCB" | elem == "RB" | elem == "RCB" | elem == "LWB" | elem == #"RWB"){
#        pos_vec <- c(pos_vec, "B")
#    } else if (elem == "CDM" | elem == "CM" | elem == "LCM" | elem == "LDM" | elem == "LM" |  elem == "LW" | elem == #"RCM" | elem == "RDM" | elem == "RW" | elem == "RM") {
#        pos_vec <- c(pos_vec, "M")
#    } else if (elem == "CAM" | elem == "CF" | elem == "LAM" | elem == "LF" | elem == "LS" |  elem == "RAM" | elem == #"RF" | elem == "RS" | elem == "ST") {
#        pos_vec <- c(pos_vec, "F")
#    } else {
#        pos_vec <- c(pos_vec, "GK")
#    }
#}

sum(is.na(test_data$Club))
sum(is.na(test_data$Age))
sum(is.na(test_data$Overall))
#sum(is.na(test_data$Club))

teams <- player_data$Club





inverseResponsePlot(predi)

#test_data$Club <- ifelse(is.na(test_data$Club), "MISSING", test_data$Club)
#test_data$Overall <- ifelse(is.na(test_data$Overall), median(test_data$Overall), test_data$Overall)
#test_data$Age <- ifelse(is.na(test_data$Age), median((test_data$Age)), test_data$Age)

#test_data <- na.omit(test_data)
#sum(is.na(test_data))

#summary(predi)
#anova(predi)
#test_data$new_pos <- pos_vec

#plot(table(player_data$Age))
#plot(table(player_data$Overall))

pred_wages <- predict(predi, test_data)
#pred_wages1 <- predict(predi, player_data)

#min_wage <- min(pred_wages[pred_wages > 0])


#for(i in 1:length(pred_wages)) {
#    print(i)
#    if(pred_wages[i] < 0) {
#        pred_wages[i] <- min_wage
#    }
#}

pred_wages <- exp(pred_wages)

#plot(pred_wages1)
#plot(player_data$WageNew)
result_matrix <- matrix(test_data$Ob)
result_matrix <- cbind(result_matrix, pred_wages)
colnames(result_matrix) <- c("Ob", "WageNew")
write.csv(result_matrix, "Test Regression", row.names = FALSE)
#nrow(result_matrix)

#table(player_data$Club)
#sum(is.na(test_data))
#pred_wages < 0



```
```{r change_club}

clubNames = unique(player_data[,"Club"])
avgSalaries <- c()

#player_data$avg <- c()

length(unique(player_data$Club))

for(club in clubNames){
    avg <- mean(player_data[which(player_data$Club == club),]$WageNew)
    player_data[which(player_data$Club == club),]$avg <- avg
}

head(player_data)
  
```


```{r pred}
library("car")
#player_data$

#player_data <- read.csv("../fifa2019wages/FifaTrainNew.csv", stringsAsFactors = FALSE)
head(player_data)
dim(player_data)

for(i in 1:length(player_data$Club)) {
    if(is.na(player_data$Club[i]) | player_data$Club[i] == "NA"){
        player_data$Club[i] <- "MISSING"
    }
}

for(i in 1:length(player_data$International.Reputation)) {
    if(is.na(player_data$International.Reputation[i])){
        player_data$International.Reputation[i] <- 1
    }
}


sum(is.na(player_data$Overall))
sum(is.na(player_data$International.Reputation))
sum(is.na(player_data$Age))
sum(is.na(player_data$Club))

#test_data <- read.csv("FifaNoY.csv", stringsAsFactors = FALSE)

test_data$Reactions[is.na(test_data$Reactions)] <- 55
test_data$Composure[is.na(test_data$Composure)] <- 55
test_data$International.Reputation[is.na(test_data$International.Reputation)] <- 1


sum(is.na(test_data$Reactions))

for(i in 1:length(test_data$Club)) {
    if(is.na(test_data$Club[i]) ){
        test_data$Club[i] <- "MISSING"
    }
    if(is.na(test_data$Overall[i])){
        test_data$Overall[i] <- median(test_data$Overall[i])
    }
    if(is.na(test_data$Age[i])){
        test_data$Age[i] <- median(test_data$Age[i])
    }
    if(is.na(test_data$International.Reputation[i])){
        test_data$International.Reputation[i] <- 1
    }
    if(is.na(test_data$ShortPassing[i])){
        test_data$ShortPassing[i] <- 55
    }
    if(is.na(test_data$BallControl[i])){
        test_data$BallControl[i] <- 55
    }
}

#predi <- lm(log(WageNew) ~ Overall + Age + International.Reputation + Composure + Special + Club + newPos + ShortPassing, data = player_data)

#predi <- lm(log(WageNew) ~ Overall + Age + International.Reputation + Composure + Special + Club, data = player_data)

#plot(predi)
#summary(predi)
#anova(predi)

sum(is.na(test_data$Club))
sum(is.na(test_data$Age))
sum(is.na(test_data$Overall))
sum(is.na(test_data$Club))
sum(is.na(test_data$Reactions))
sum(is.na(test_data$Composure))
sum(is.na(test_data$Special))
sum(is.na(test_data$ShortPassing))
sum(is.na(test_data$BallControl))



vif(predi)

cor(player_data$Age, player_data$Overall)

inverseResponsePlot(predi)

pred_wages <- predict(predi, test_data)

pred_wages <- exp(pred_wages)


result_matrix <- matrix(test_data$Ob)
result_matrix <- cbind(result_matrix, pred_wages)
colnames(result_matrix) <- c("Ob", "WageNew")
write.csv(result_matrix, "Test Regression", row.names = FALSE)

plot(pred_wages)
```

```{r test}
library(ggplot2)

pos_vec <- c()
for(elem in player_data$Position) {
   if (is.na(elem)) {
       pos_vec <- c(pos_vec, "0")
   } else if(elem == "CB" | elem == "LB" | elem == "LCB" | elem == "RB" | elem == "RCB" | elem == "LWB" | elem == "RWB"){
       pos_vec <- c(pos_vec, "B")
   } else if (elem == "CDM" | elem == "CM" | elem == "LCM" | elem == "LDM" | elem == "LM" |  elem == "LW" | elem == "RCM" | elem == "RDM" | elem == "RW" | elem == "RM") {
       pos_vec <- c(pos_vec, "M")
   } else if (elem == "CAM" | elem == "CF" | elem == "LAM" | elem == "LF" | elem == "LS" |  elem == "RAM" | elem == "RF" | elem == "RS" | elem == "ST") {
       pos_vec <- c(pos_vec, "F")
   } else {
       pos_vec <- c(pos_vec, "GK")
   }
}

player_data$newPos = pos_vec

ggplot(data = player_data, aes(x = newPos, y = WageNew)) + geom_point()

?plot
#plot(predi)
```


```{r step}

library(MASS)
# Fit the full model 
full.model <- lm(WageNew ~ Age + Overall + Potential + Special + International.Reputation + Composure + Weak.Foot + Skill.Moves + newPos, data = player_data)
# Stepwise regression model


step.model <- stepAIC(full.model, direction = "forward", 
                      trace = FALSE, k = log(length(player_data$WageNew)))
summary(step.model)

step.model$anova


```








 THIS IS WHERE HW5 STARTS


```{r teams}
library(ggplot2)
#player_data$Joined
dim(player_data)
summary(player_data$WageNew)

#ggplot(data = player_data, aes(x = Joined, y = WageNew)) + geom_point()
#for i in seq_len(len(player_data$Club)):
 # player_data$Club[i]

```


## Problem 1

### a) Group Name: Kaiqi Zhu, Lec 1D
Teammates Names: Yonatan Khalil, Matei Lupu, Jenna Park, Kaiqi Zhu

### b) 
We kept the NAs and replaced them with values that would help us analyze the data set accurately. For example, when a player was missing a skill value we gave them one on the lower end  as they tended to be players with low overall values. This meant that our dataset still had 12745 rows and 81 columns.

### c)
The following is the summary for the response variable.             
```{r Part1}

player_data <- read.csv("fifa2019wages/FifaTrainNew.csv", stringsAsFactors = FALSE)
summary(player_data$WageNew)

```

### d)
We have 6 predictors for our most recent prediction.

### e)
```{r pred}

pred <- matrix(c("Overall", "Age", "International.Reputation", "Composure", "Special", "ShortPassing", "Club", "newPos", NA, NA, NA, NA), ncol = 2)
colnames(pred) = c("Numerical", "Categorical")

head(pred)
```

### f)
Our latest rank on Kaggleis 11 and our R^2 is 0.95807.

### g)
```{r plt}
data <- player_data[, c("WageNew", "Overall", "Age", "International.Reputation", "Composure", "Special", "ShortPassing", "newPos")]
plot(data)
```

### h)
```{r cor}
library(corrplot)

pred_data <- player_data[, c("WageNew", "Overall", "Age", "International.Reputation", "Composure", "Special", "ShortPassing",)]
cor_data <- na.omit(pred_data)
vals <- cor(cor_data[sapply(cor_data, is.numeric)])

corrplot(vals)
```


## Problem 5

```{r best}
library(leaps)
sub_data = na.omit(player_data)
sub_data1 <-  sub_data[, c("Overall", "Age", "International.Reputation", "Composure", "Special", "ShortPassing")]

sub <- regsubsets(as.matrix(sub_data1), log(sub_data$WageNew)) 
sub_sum <- summary(sub)

par(mfrow=c(1,1))
plot(1:6,sub_sum$adjr2,xlab="Subset Size",ylab="Adjusted R-squared")

om1 <- lm(log(WageNew) ~ Overall, data = sub_data)
om2 <- lm(log(WageNew) ~ Overall + Age, data = sub_data)
om3 <- lm(log(WageNew) ~ Overall + Age + International.Reputation, data = sub_data)
om4 <- lm(log(WageNew) ~ Overall + Age + International.Reputation + Composure, data = sub_data)
om5 <- lm(log(WageNew) ~ Overall + Age + International.Reputation + Composure + Special, data = sub_data)
om6 <- lm(log(WageNew) ~ Overall + Age + International.Reputation + Composure + Special + ShortPassing, data = sub_data)

om7 <- predi

n <- length(om1$residuals) 
extractAIC(om1, k = 2)
extractAIC(om1, k = log(n))

n <- length(om2$residuals) 
extractAIC(om2, k = 2)
extractAIC(om2, k = log(n))

n <- length(om3$residuals) 
extractAIC(om3, k = 2)
extractAIC(om3, k = log(n))

n <- length(om4$residuals) 
extractAIC(om4, k = 2)
extractAIC(om4, k = log(n))

n <- length(om5$residuals) 
extractAIC(om5, k = 2)
extractAIC(om5, k = log(n))

n <- length(om6$residuals) 
extractAIC(om6, k = 2)
extractAIC(om6, k = log(n))

n <- length(om7$residuals) 
extractAIC(om7, k = 2)
extractAIC(om7, k = log(n))


```


```{r P5b}
resp <- lm(log(WageNew)~1, data=player_data)

backwardAIC <- step(resp,scope=list(lower=~1,
upper=~Overall + Age + International.Reputation + Composure + Special + newPos + ShortPassing,
direction="backward", data=player_data, k = 2))

backwardBIC <- step(resp, scope=list(lower=~1,
upper=~Overall + Age + International.Reputation + Composure + Special + newPos + ShortPassing,
direction="backward", data=player_data, k=log(nrow(player_data))))


```


```{r P5c}
resp <- lm(log(WageNew)~1, data=player_data)

forwardAIC <- step(resp,scope=list(lower=~1,
upper=~Overall + Age + International.Reputation + Composure + Special + newPos + ShortPassing,
direction="forward", data=player_data, k = 2))

forwardBIC <- step(resp, scope=list(lower=~1,
upper=~Overall + Age + International.Reputation + Composure + Special + newPos + ShortPassing,
direction="forward", data=player_data, k=log(nrow(player_data))))


```

